---
title: "Prepare meta-analysis dataset"
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "3.1-prepare-meta-analysis-dataset.html"
    # embed-resources: true
    # code-fold: true
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---

Now we'll meta-analyze the quantitative articles that provided data usable for a meta-analysis. We'll start by calculating effect sizes and then converting effect sizes to a common type for analysis based on our analysis groups. 

We'll then run a suite of models and choose the model with the best random effects. Finally, we'll plot the resulting model estimates and conduct publication bias tests.

# Prepare dataset
## Prepare environment
We'll start by loading packages and some helper functions. The `convert_effect_sizes` function contains formulas to convert between effect sizes. The formulas were originally provided by @borenstein2021. The `get_r` function is a simple helper function for extracting the correlation coefficient (`r`) from the `cor.test` output.

```{r}
rm(list = ls())

library("data.table")
library("metafor")
library("ggplot2")
library("ggstance")
library("orchaRd")
library("patchwork")
library("readxl")
library("stringr")
library("sf")
library("mapview")
library("dplyr")
library("rotl")
library("broom.mixed")
library("here")

source("scripts/functions/convert_effect_sizes.R") # This function converts between effect sizes

get_r <- function(x, y){
  out <- cor.test(x, y)
  return(out$estimate)
}

```

## Calculate effect sizes {#sec-calc-effect-sizes}
We'll start by loading the raw extracted data. Given the heterogeneity in experimental designs, the raw data was collected in 3 separate spreadsheets based on the data structure required for each effect size type. We'll load these into a list object:

```{r}
#| eval: true
#| echo: true
#| output: false

files <- c("data/meta_analysis/Cat meta-analysis - SMDs.xlsx",
           "data/meta_analysis/Cat meta-analysis ORs.xlsx",
           "data/meta_analysis/Cat meta-analysis - correlations.xlsx")
dat_list <- lapply(files, read_excel)
names(dat_list) <- c("SMDs", "ORs", "Zrs")

```

We also have a separate file that gives analysis groups for each effect size---e.g., how to group effect sizes for each model. These groupings are also the scaffold for some of our sensitivity analyses (e.g., different degrees of lumping across different experimental treatment types).

The analysis groups include:

- Abundance with/without cats
- Abundance before-after eradication
- Abundance on islands with/without cats
- Abundance spatial association
- Abundance temporal association
- Reproduction with/without cats
- Reproduction before-after eradication
- Reproduction temporal association

The 'with/without cats' is a broader analysis group that includes the data in 'before-after eradication' and 'islands with/without cats'. 

```{r}
#| eval: true
#| echo: true

analysis_groups <- read_excel("data/meta_analysis/analysis groups.xlsx",
                              skip = 0) |> setDT()

unique(analysis_groups[, .(Description, final_analysis_group1, final_analysis_group2)])

```


### z-to-r-transformed correlation coefficients (Zr) effect sizes {#sec-calc-Zr}
We'll start with the correlation coefficient (Zr) effect sizes. 

Note that several effect sizes in this dataset can be analyzed either as Zr effect sizes (with binary predictor) or SMD. We've decided that they would best be analyzed as SMD effect sizes, which we handle in the code below.

Because the variance of proportion data increases closer to 0/1 boundaries, we stabilized variance with arc-sin square root transformation and the delta method following @Macartney2022-ik.

```{r}
#| eval: true
#| echo: true
#| output: false

cor <- dat_list[["Zrs"]] |> setDT()

sort(unique(cor$Effect_size_ID))

# >>> Clean -------------------------------------------------------------------
unique(cor$Description)

cor[Description == "Spatial correlation" & Effect_size_ID == "ES_3", ]

# Make variables numeric:
cor[is.na(as.numeric(Prey_correlation_data))]$Prey_correlation_data
cor[, Prey_correlation_data := str_trim(Prey_correlation_data)]
cor[is.na(as.numeric(Prey_correlation_data))]$Prey_correlation_data

cor[, Prey_correlation_data := as.numeric(Prey_correlation_data)]
cor[is.na(as.numeric(Cat_correlation_data))]$Cat_correlation_data
cor[, Cat_correlation_data := as.numeric(Cat_correlation_data)]

class(cor$cat_presence)
cor[is.na(as.numeric(cat_presence))]$cat_presence

# So these are binary predictor. ALl of these except one can be SMD
# The other can be Zr with 0/1 predictor.
cor[Effect_size_ID == "ES_47b", Cat_correlation_data := cat_presence]
cor[Effect_size_ID == "ES_47b"]

cor.smd <- cor[is.na(Cat_correlation_data), ]
cor.smd$cat_presence
cor.smd[, .(n_points = .N), by = .(cat_presence, Effect_size_ID)]

# >>> Calculate ZCOR --------------------------------------------------
(cor)

cor.sum <- cor[!is.na(Cat_correlation_data), 
               .(r = get_r(x = Cat_correlation_data,
                           y = Prey_correlation_data),
                   n = .N),
               by = .(Effect_size_ID, scientificName,  Description,
                      unit_of_replication,
                      Article, Article_secondary_same_data, 
                      analysis_group, effect_size_type, Abundance_reproduction,
                      Prey_units, Cat_units, Source, Analysis_comment, 
                      Latitude, Longitude, Study_location)]

cor.sum[duplicated(Effect_size_ID)]
cor.sum[Effect_size_ID == "ES_32", 
        Effect_size_ID := paste0(Effect_size_ID, "_", .GRP), 
        by = .(scientificName)]
cor.sum[Effect_size_ID == "ES_47b"]
cor[Effect_size_ID == "ES_47b"]

cor.sum <- escalc(measure = "ZCOR",
                     ri = r, ni = n,
                     data = cor.sum) |> setDT()

length(unique(cor.sum$Effect_size_ID))
length(unique(cor.sum$Article))

names(cor.sum)

unique(cor.sum$analysis_group)

cor.sum[, original_effect_size := "Zr"]

# > Convert groups that can actually be SMD --------------------------------------------
unique(cor.smd$analysis_group)

cor.smd[analysis_group == "Short-term reproduction", .(n = .N), by = .(cat_presence)]

cor.smd <- cor.smd[, .(mean_prey = mean(Prey_correlation_data),
                       sd_prey = sd(Prey_correlation_data),
                       n = .N),
                   by = .(Effect_size_ID, scientificName, Article, Article_secondary_same_data,
                          analysis_group,effect_size_type, Description,
                          unit_of_replication,
                          Abundance_reproduction, Prey_units, Cat_units, Source,
                          Latitude, Longitude, Study_location,
                          cat_presence
                          )]

# Transform 0-1 bounded data following Macartney et al 2022 Neurosci Biobehav Rev:
unique(cor.smd$Effect_size_ID)
cor.smd[Effect_size_ID %in% c("ES_43", "ES_37"),]

cor.smd[Effect_size_ID %in% c("ES_43"), `:=` (sd_prey = sd_prey/100,
                                              mean_prey = mean_prey/100)]
cor.smd[Effect_size_ID %in% c("ES_43", "ES_37"), 
        `:=` (sd_prey = sqrt( sd_prey^2 / ((4*mean_prey) * (1-mean_prey)) ))]
cor.smd[Effect_size_ID %in% c("ES_43", "ES_37"), mean_prey := asin(sqrt(mean_prey))]
cor.smd

# Make wide by cat presence/absence
cor.smd.wide <- dcast(cor.smd,
                     ... ~ cat_presence,
                     value.var = c("mean_prey", "sd_prey", "n"))

cor.smd.wide
setnames(cor.smd.wide, c("n_1", "n_0"), c("n1", "n2"))

cor.smd.wide
cor.smd.out <- escalc(measure = "SMD", 
                 m1i = mean_prey_1, m2i = mean_prey_0,
                 sd1i = sd_prey_1, sd2i = sd_prey_1,
                 n1i = n1, n2i = n2,
                 data = cor.smd.wide) |> 
  setDT()
cor.smd.out[, n := n1 + n2] # for conversions, later

cor.smd.out
cor.smd.out[, original_effect_size := "SMD"]

# >>> Bind ----------------------------------------------------------------

cor.sum
cor.final <- rbind(cor.sum, 
                   cor.smd.out,
                   fill = T)

cor.final[duplicated(Effect_size_ID)]

# Must be 0 rows
cor.final[Effect_size_ID == "ES_37", ]
cor.final[Effect_size_ID == "ES_37", ]

# This one is in SMD dataset:
cor.final <- cor.final[Effect_size_ID != "ES_ARM", ]

```

### Standardized Mean Difference (SMD) effect sizes {#sec-calc-smd}

Now let's calculate SMD for the effect sizes in the SMD file. This data reflects two discrete treatments (cats present, cats absent) with a continuous response variable summarized as mean±SD.

```{r}
#| eval: true
#| echo: true
#| output: false

smd <- dat_list[["SMDs"]] |> setDT()
head(smd)

# >>> Clean -------------------------------------------------------------------
# setnames(smd, "...30", "analysis_group_notes")
smd[is.na(as.numeric(Prey_mean_cats_Absent))]
smd[is.na(as.numeric(Prey_mean_cats_Present))]
smd[is.na(as.numeric(Prey_error_cats_Absent))]
smd[is.na(as.numeric(Prey_error_cats_Present))]
smd[is.na(as.numeric(Sample_size_overall_cats_Absent))]
smd[is.na(as.numeric(Sample_size_overall_cats_Present))]
smd[, `:=` (Prey_mean_cats_Absent = as.numeric(Prey_mean_cats_Absent),
            Prey_mean_cats_Present = as.numeric(Prey_mean_cats_Present),
            Prey_error_cats_Absent = as.numeric(Prey_error_cats_Absent),
            Prey_error_cats_Present = as.numeric(Prey_error_cats_Present),
            Sample_size_overall_cats_Absent = as.numeric(Sample_size_overall_cats_Absent),
            Sample_size_overall_cats_Present = as.numeric(Sample_size_overall_cats_Present))]

unique(smd$Effect_size_ID)

# >>> Calculate effect sizes --------------------------------------------------
setnames(smd, c("Sample_size_overall_cats_Absent", "Sample_size_overall_cats_Present"),
                c("n2", "n1")) # This is to prevent later confusion while converting

smd <- escalc(measure = "SMD",
              m2i = Prey_mean_cats_Absent, m1i = Prey_mean_cats_Present,
              sd2i = Prey_error_cats_Absent, sd1i = Prey_error_cats_Present,
              n2i = n2, n1i = n1,
              data = smd) |> setDT()
smd[, .(Prey_mean_cats_Absent, Prey_mean_cats_Present, yi)]

smd[, original_effect_size := "SMD"]

# smd <- rbind(smd)
smd$analysis_group

# smd[converted_to == "Zr"]
smd.final <- copy(smd)

```

### log Odds Ratio (lnOR) effect sizes {#sec-calc-lnOR}

Now we'll calculate our odds ratio effect. This data has two discrete predictors (cats present / absent) and a binary outcome variable (present versus absent or nest fledged versus failed).

```{r}
#| eval: true
#| echo: true
#| output: false

or <- dat_list[["ORs"]] |> setDT()
or

#
or <- escalc(measure = "OR",
             ai = Prey_and_cat_positive,
             bi = Prey_negative_cat_positive,
             ci = Prey_positive_cat_negative,
             di = Prey_and_cat_negative,
             data = or) |> setDT()
or

or[, n1 := Prey_and_cat_positive + Prey_negative_cat_positive]
or[, n2 := Prey_positive_cat_negative + Prey_and_cat_negative]
or[, original_effect_size := "lnOR"]
or.final <- copy(or)

```

### Bind effect size datasets together and prepare analysis groups

Now we bind these datasets together:
```{r}
#| eval: true
#| echo: true
#| output: true

meta_combined <- rbind(cor.final, smd.final, or.final, fill = T)

# Make sure there are no duplicated effect sizes
meta_combined[duplicated(Effect_size_ID)]

```

Now we're going to merge in the analysis groups, for which there are multiple effect size ID. This merge will increase the number of rows of the dataset (but each effect size will only be in each model once because the analysis group variable will be unique for each model).

```{r}
#| eval: true
#| echo: true
#| output: false

meta_combined[, key := paste(Abundance_reproduction, Description)]
analysis_groups[, key := paste(Abundance_reproduction, Description)]

analysis_groups <- melt(analysis_groups,
                        id.vars = c("key"),
                        measure.vars = c("final_analysis_group1", "final_analysis_group2"),
                        variable.name = "analysis_group_level",
                        value.name = "final_analysis_group")
analysis_groups <- analysis_groups[!is.na(final_analysis_group)]
analysis_groups[, analysis_group_level := ifelse(analysis_group_level=="final_analysis_group1",
                                                 "fine_scale", "lumped")]

setdiff(meta_combined$key, analysis_groups$key)
setdiff(analysis_groups$key, meta_combined$key)

analysis_groups[, .(n = uniqueN(final_analysis_group)), by = .(key)]
# 

meta_combined.mrg <- merge(meta_combined[, !c("analysis_group")],
                           unique(analysis_groups[, .(key, analysis_group_level, final_analysis_group)]),
                           by = "key",
                           allow.cartesian = TRUE,
                           all.x = T)

meta_combined.mrg[is.na(final_analysis_group)]

meta_combined.mrg[, n_effect_types := uniqueN(original_effect_size),
                  by = .(final_analysis_group)]
meta_combined.mrg[n_effect_types == 1, ]

setnames(meta_combined.mrg, "final_analysis_group", "analysis_group")

```

Let's check that the units of replication are not too different per analysis group:
```{r}
meta_combined.mrg[, .(n_units = uniqueN(unit_of_replication),
                      units = paste(unique(unit_of_replication), collapse = "; ")),
                  by = .(analysis_group)]
```

We think that these are close enough to be acceptable. What we really don't want are analysis groups that mix individuals (or nests) with surveys/plots/islands/sites.

## Convert effect sizes {#sec-convert-effect-sizes}
Now we'll begin to convert effect sizes. We'll start by splitting out analysis groups that are a single effect size type (i.e., don't require conversion).

```{r}
#| eval: true
#| echo: true
one_type <- meta_combined.mrg[n_effect_types == 1, ]

one_type[n_effect_types == 1, `:=` (yi_analysis = yi,
                                    vi_analysis = vi,
                                    analysis_effect_size = original_effect_size)]
nrow(one_type)
length(unique(one_type$analysis_group))
```

Now we'll deal with mixed groups. First we'll determine what the dominant group is and use this to designate the analysis effect size for each analysis group. 

```{r}
#| eval: true
#| echo: true
#| output: false

mixed_types <- meta_combined.mrg[n_effect_types > 1, ]

# Determine dominant effect size type per analysis group
mixed_types[, .(paste(sort(unique(original_effect_size)), 
                      collapse = "; ")),
                by = .(analysis_group)]

type_freq <- mixed_types[, .(n = .N),
                 by = .(analysis_group, original_effect_size)]
setorder(type_freq, analysis_group)

type_freq <- type_freq[, .SD[which.max(n)], by = .(analysis_group)]

setnames(type_freq, "original_effect_size", "analysis_effect_size")

mixed_types <- merge(mixed_types,
                     type_freq[, .(analysis_group, analysis_effect_size)], 
                     by = "analysis_group",
                     all.x = T)

mixed_types[analysis_group %in% c( "Reproduction spatial association"),
            analysis_effect_size := "SMD"]

mixed_types[is.na(analysis_effect_size), ]

already_converted <- mixed_types[analysis_effect_size == original_effect_size, ]
to_convert <- mixed_types[analysis_effect_size != original_effect_size, ]

#
to_convert[, key := paste(original_effect_size, analysis_effect_size)]

```

We'll now loop over this dataset based on the unique combinations of original effect size (`original_effect_size`) and destination effect size (`analysis_effect_size`). The effect size formulas we use here are given [here](3.4-effect-size-conversions.qmd).

```{r}
#| eval: true
#| echo: true
#| output: false

combos <- unique(to_convert$key)
combos
out <- list()

convert_effect_sizes(formula_path = "scripts/functions/conversion_formulas.csv")

for(i in 1:length(combos)){
  sub_dat <- to_convert[key == combos[i]]
  
  out[[i]] <- convert_effect_sizes(n = n, 
                                   n1 = n1, n2 = n2,
                                   yi = yi, vi = vi, 
                                   r = r,
                                   from = unique(sub_dat$original_effect_size),
                                   to = unique(sub_dat$analysis_effect_size),
                                   data = sub_dat,
                                   bind = TRUE,
                                   formula_path = "scripts/functions/conversion_formulas.csv")
}

converted <- rbindlist(out)
setnames(converted, c("yi_trans", "vi_trans"), c("yi_analysis", "vi_analysis"))
already_converted[, `:=` (yi_analysis = yi, vi_analysis = vi)]

```
Now we'll bind the dataset back together:

```{r}
#| eval: true
#| echo: true
#| output: true

mixed_type <- rbind(converted, already_converted)
mixed_type[is.na(yi_analysis), ]

meta.final <- rbind(one_type, mixed_type)
```


## Merge in covariates {#sec-merge-in-covariates}

### Body mass
Now we'll merge in covariates. We'll start with body mass. Body mass for mammals was derived from PHYLACINE [Faurby2018-wu] and is available for download [here](https://github.com/MegaPast2Future/PHYLACINE_1.2). Body mass for birds was derived from Avonet [@Tobias2022-eu] and is available for download [here](https://figshare.com/s/b990722d72a26b5bfead). Body mass for reptiles and species absent from the other datasets include [@San-Diego-Zoo-Wildlife-Alliance-Library2015-qi; @Iucn2018-qz; @Gibbs2001-st].

```{r}
#| eval: true
#| echo: true
#| output: false

species <- data.table(scientificName = unique(c(meta.final$scientificName)))

species[, spp_name_corrected := scientificName]
species[scientificName == "Pampusana erythroptera", 
        spp_name_corrected := "Gallicolumba erythroptera"]

# Mammals:
phyl <- fread("data/trait_databases/phylacine_traits.csv")

phyl[, Binomial.1.2 := gsub("_", " ", Binomial.1.2)]

species.m1 <- merge(species, phyl[, .(Binomial.1.2, Order.1.2, 
                                      Mass.g)],
                    by.x = "spp_name_corrected", by.y = "Binomial.1.2",
                    all.x = T, all.y = F)

# Birds:
avonet <- read_excel("data/trait_databases/AVONET Supplementary dataset 1.xlsx",
                     "AVONET1_BirdLife") |> setDT()
unique(avonet$Species1)

species.m1[scientificName %in% avonet$Species1]

species.m2 <- merge(species.m1,
                    avonet[, .(Species1, Order1, Mass)],
                    by.x = "spp_name_corrected", by.y = "Species1",
                    all.x = T, all.y = F)

species.m2

species.m2[, Mass_g_final := ifelse(is.na(Mass.g), Mass, Mass.g)]
species.m2[, Order_final := ifelse(is.na(Order.1.2), Order1, Order.1.2)]
species.m2[is.na(Mass_g_final)]

species.m2[, source := fcase(!is.na(Order.1.2), "Phylacine 1.2",
                             !is.na(Order1), "Avonet",
                             is.na(Order.1.2) & is.na(Order1), NA)]

species.m2[, class := fcase(!is.na(Order.1.2), "Mammals",
                             !is.na(Order1), "Birds",
                             is.na(Order.1.2) & is.na(Order1), NA)]


species.m2[is.na(Mass_g_final)]
species.m2[spp_name_corrected == "Chelonoidis niger", `:=` (Order_final = "Testudines",
                                                            Mass_g_final = 80,# weight at birth
                                                            class = "Reptiles",
                                                            source = "IUCN Red List; https://ielc.libguides.com/sdzg/factsheets/galapagostortoises/summary")]
species.m2[spp_name_corrected == "Cyclura carinata", `:=` (Order_final = "Squamata",
                                                            Mass_g_final = 355,
                                                           class = "Reptiles",
                                                            source = "IUCN Red List;  ⁠Alberts, A. (1999) West Indian Iguanas: Status Survey and Conservation Action Plan. IUCN/SSC West Indian Iguana Specialist Group. IUCN, Gland, Switzerland and Cambridge ")]

species.m2[spp_name_corrected == "Gallicolumba erythroptera", `:=` (Order_final = "Columbiformes",
                                                           Mass_g_final = 113.5,
                                                           class = "Birds",
                                                           source = "Gibbs, David; Barnes, Eustace; Cox, John (2001). Pigeons and Doves: A Guide to the Pigeons and Doves of the World. Sussex: Pica Press. ISBN 1-873403-60-7.")]

species.m2[is.na(class), ]

meta.final.mrg <- merge(meta.final,
                           species.m2[, .(scientificName, Order_final, Mass_g_final, class)],
                           by = "scientificName",
                           all.x = T)

nrow(meta.final.mrg) == nrow(meta.final)
# Must be TRUE

meta.final.mrg[is.na(class), ]
meta.final.mrg[scientificName == "Prosobonia cancellata",
               `:=` (class = "Birds",
                     Mass_g_final = 33)]
meta.final.mrg[class == "", ]
meta.final.mrg
unique(meta.final.mrg$class)

# Now let's log the body mass:
meta.final.mrg[, log_mass := log10(Mass_g_final)]

```

We'll also be sure to filter out species that we are excluding (because cat impacts were claimed to be via competition not predation, which is outside the scope of our study).

```{r}
#| eval: true
#| echo: true
#| output: true

sort(unique(meta.final.mrg$scientificName))
meta.final.mrg <- meta.final.mrg[!scientificName %in% c("Dasyurus maculatus", 
                                                        "Fossa fossana",
                                                        "Cryptoprocta ferox")]

nrow(meta.final.mrg)

```

### Habitat use
Now we'll merge in habitat classifications:

```{r}
#| eval: true
#| echo: true
#| output: false

habitat <- read_excel("data/meta_analysis/Habitat_use.xlsx") |> setDT()
head(habitat)

names(habitat)
habitat <- habitat[, .(scientificName, locomotion_volant,
                       ground_or_burrow_resting_or_nesting, 
                       Foraging_habitat_ground)]
habitat[, `:=` (locomotion_volant = ifelse(locomotion_volant == 0, 
                                           "non_volant", "volant"))]
habitat[, `:=` (ground_or_burrow_resting_or_nesting = 
                  ifelse(ground_or_burrow_resting_or_nesting == 0, 
                         "ground", 
                         "other"))]
habitat[, `:=` (Foraging_habitat_ground = ifelse(Foraging_habitat_ground == 0, 
                                                 "ground", 
                                                 "other"))]
habitat

setdiff(meta.final.mrg$scientificName, habitat$scientificName)

meta.final.mrg2 <- merge(meta.final.mrg,
                        habitat,
                        by = "scientificName")

nrow(meta.final.mrg2) == nrow(meta.final.mrg)
# Must be TRUE
```

### Inside prey range
Now let's designate whether they are inside of cat prey range based on @Wooster2024-gc:

```{r}
#| eval: true
#| echo: true
#| output: true

meta.final.mrg2[, prey_range_primary := ifelse(Mass_g_final >= 17 & 
                                                Mass_g_final <= 2000, "inside", "outside")]
meta.final.mrg2[, prey_range_secondary := ifelse(Mass_g_final >= 8 & 
                                                  Mass_g_final <= 2000, "inside", "outside")]
meta.final.mrg2[, prey_range_tertiary := ifelse(Mass_g_final >= 6 & 
                                                 Mass_g_final <= 2650, "inside", "outside")]
nrow(meta.final.mrg2)

```
### Continents versus islands
Now we'll use the study latitude and longitude to determine if it occurred on a continent or island. Note that the only continental studies found come from Australia.

```{r}
#| eval: true
#| echo: true
#| output: false

continents <- st_read("data/spatial/4a7d27e1-84a3-4d6a-b4c2-6b6919f3cf4b202034-1-2zg7ul.ht5ut.shp")
# The largest of each is the continent and the rest are islands...

# Split into single part polygons and identify the largest polygon of each continent:
continents <- continents %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(.)) %>%
  arrange(CONTINENT, -area) %>%
  group_by(CONTINENT) %>%
  mutate(continent_island = c("mainland", rep("island", (n()-1))))
continents

continents %>%
  filter(continent_island == "mainland") %>%
  mapview()

continents %>%
  filter(continent_island != "mainland") %>%
  mapview()

continents %>% filter(CONTINENT == "Oceania")
continents <- continents %>%
  mutate(continent_island = ifelse(CONTINENT == "Oceania", 
                                   "island", 
                                   continent_island))

# A few other places aren't really islands...the UK for instance...
continents <- continents %>%
  mutate(poly_id = seq(1:n()))

continents %>%
  filter(continent_island != "mainland") %>%
  mapview()

continents <- continents %>%
  mutate(continent_island = ifelse(poly_id %in% c(),
                                   "mainland", continent_island))

meta.final.mrg$Longitude
meta.final.mrg[is.na(as.numeric(Longitude))]
meta.final.mrg[is.na(as.numeric(Latitude))]

meta.spat <- meta.final.mrg %>%
  unique() %>%
  select(Effect_size_ID, Longitude, Latitude) %>%
  st_as_sf(coords = c("Longitude", "Latitude"),
           crs = 4326)
meta.spat

meta.spat.jnd <- st_join(meta.spat,
                         continents %>% select(continent_island))
nrow(meta.spat.jnd) == nrow(meta.spat)

meta.spat.jnd
meta.spat.jnd %>%
  filter(is.na(continent_island)) %>% 
  mapview(zcol = "continent_island")
# ALl of these NAs are on islands

meta.spat.jnd <- meta.spat.jnd %>%
  mutate(continent_island = ifelse(is.na(continent_island), "island", continent_island))

meta.spat.jnd

meta.spat.jnd %>%
  filter(continent_island == "island") %>% 
  mapview(zcol = "continent_island")

meta.spat.jnd %>%
  filter(continent_island == "mainland") %>% 
  mapview(zcol = "continent_island")
# Only australia

meta.spat.jnd <- meta.spat.jnd %>%
  as.data.frame() %>%
  mutate(geometry = NULL) %>% 
  setDT() %>%
  unique()
meta.spat.jnd

meta.spat.jnd[duplicated(Effect_size_ID), ]

nrow(meta.final.mrg)
meta.final.mrg3 <- merge(meta.final.mrg2,
                        meta.spat.jnd,
                        by = "Effect_size_ID")

nrow(meta.final.mrg3) == nrow(meta.final.mrg2)
# Must be TRUE
```

### Final checks
Now we'll do some final checks before proceding to analysis

```{r}
#| eval: true
#| echo: true
#| output: true

meta.final.mrg3[, .(n_effects = uniqueN(analysis_effect_size)), by = .(analysis_group)]
# All should be 1

# >>> Check for duplicates by ID ------------------------------------------
meta.final.mrg3[duplicated(paste(Effect_size_ID, analysis_group))]
# Must be 0 rows

# Save final meta-analysis dataset
fwrite(meta.final.mrg3, "builds/meta_analysis/analysis_ready_dataset.csv")

```
