---
title: "Fit meta-analysis models"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "3.2-run-models.html"
    # embed-resources: true
    code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---

Now we'll fit multilevel meta-analytic models in the R package `metafor` [@Viechtbauer2010-uq]. This workflow is derived from a full R script in `scripts/2_analysis/2 fit meta analysis models.R` that includes additional checks and tests and is fully interactive if you clone the repo.

Given the non-independence structures in this data we will fit a series of models with increasingly complex random effects, particularly to control for phylogenetic similarity between prey species and to also deal with non-phylogenetic components of species identity. We'll create a model guide to organize these models, as well as our subgroup analyses and sensitivity analyses. These analyses include:

- Only prey inside of cat's prey range
- Only dominant effect sizes (e.g., excluding minority converted effect sizes)
- Prey body mass
- Habitat / locomotion / landform subgroup analyses

Note that instead of including habitat type, locomotion, or landform as categorical predictors, we used separate univariate (intercept-only) models on these subgroups of the data. In part this is to keep the code cleaner (the results are the same regardless) but also because in most cases there was only one factor level of these variables with sufficient data to estimate coefficients and we were not interested in testing for differences between factor levels.


# Prepare environment
We'll start by cleaning the environment and loading some custom functions

```{r}
#| eval: true
#| echo: true
#| output: true

# Clean environment:
rm(list = ls())

# Load packages
library("data.table")
library("metafor")
library("orchaRd")
library("readxl")
library("rotl")
library("broom.mixed")
library("stringr")
library("dplyr")

# Load custom functions:
# This helps extract predictions on rma.mv objects:
rma_predictions <- function(m, 
                            newgrid,
                            has_intercept = T){
  
  
  if(!is.data.frame(newgrid)){errorCondition("ERROR newgrid must be a data frame")}
  #create the new model matrix. 
  
  if(!all(unlist(lapply(names(newgrid), # lapply through names of newgrid to check that they're in formula
                        function(x) grepl(pattern=x,
                                          x = as.character(m$formula.mods)[-1]))))){
    errorCondition("ERROR: variables in newgrid are not in model formula")
  }
  
  
  # Drop levels that might be missing from the model...
  cols <- names(newgrid)
  coef_nms <- names(coef(m))
  temp <- c()
  
  if(has_intercept == F){
    
    for(i in 1:length(cols)){
      if(class(unlist(newgrid[, cols[i], with = F])) %in% c("factor", "character")){
        temp <- paste0(names(newgrid[, cols[i], with = F]),
                       unlist(newgrid[, cols[i], with = F]))
        newgrid <- newgrid[temp %in% coef_nms, ]
      }
    }
    
  }
  
  newgrid
  
  # Create prediction matrix
  predgrid <- (model.matrix(m$formula.mods, data=newgrid))
  predgrid
  
  if(any(grepl("intercept", colnames(predgrid), 
               ignore.case = TRUE))){
    #if intercept is present, remove it?
    predgrid <- predgrid[, -1]
  }
  
  # predict onto the new model matrix
  pred.out <- as.data.frame(predict(m, newmods=predgrid))
  
  #attach predictions to variables for plotting
  final.pred <- cbind(newgrid, pred.out)
  
  return(final.pred)
}


# This function creates a phylogenetic covariance matrix
createPhyloCorr <- function(spp_list, tree){
  
  tree.filt <- keep.tip(tree, spp_list)
  tree.br <- compute.brlen(tree.filt)
  tree.corr <- vcv(tree.br, corr=T)
  return(tree.corr)
}


```


## Prepare the dataset {#sec-prepare-data}
Let's reload our dataset. We'll clean up the article ID to be a bit friendlier as a text string in the models and we'll add a variable to our dataset to specify whether the effect size type was dominant (for filtering out minority converted effect sizes). Finally, we'll reshape our dataset to be *long* by our habitat and landform variables to make our models easier to succinctly fit:

```{r}
#| eval: true
#| echo: true
#| output: false

# Load prepared data:
dat <- fread("builds/meta_analysis/analysis_ready_dataset.csv")

# Clean up the article ID a bit:
dat[, article_id := paste(word(Article, 1, 
                               sep = "[[:space:][:punct:]]"), 
                          .GRP), 
    by = .(Article)]

# Check that that didn't screw anything up:
length(unique(dat$Article)) == length(unique(dat$article_id))

nrow(dat[duplicated(article_id)]) == nrow(dat[duplicated(Article)])
# Must be TRUE

# >>> Add column for whether effect size type is dominant within analysis group --------
dat[, total_n_articles := uniqueN(article_id), by = .(analysis_group)]
dat[, n_articles_per_original_es := uniqueN(article_id), by = .(analysis_group,
                                                                     original_effect_size)]
dat[, dominant_effect_size := ifelse((n_articles_per_original_es / total_n_articles) > .5,
                                          "yes", "no")]
dat[, .(n = .N), by = .(dominant_effect_size)]

# >>> Make data long by habitat traits ------------------------------------
# Instead of having these as separate factors, let's do univariate subgroup models
dat.long <- melt(dat,
                 measure.vars = c("locomotion_volant",
                                  "ground_or_burrow_resting_or_nesting",
                                  "Foraging_habitat_ground", "continent_island"),
                 variable.name = "habitat_category",
                 value.name = "habitat_trait")

dat[, habitat_trait := "All"]
dat[, habitat_category := "All"]

dat.final <- rbind(dat[, !c("locomotion_volant",
                            "ground_or_burrow_resting_or_nesting",
                            "Foraging_habitat_ground", "continent_island"),
                       with = F], 
                   dat.long)
head(dat.final)

# Now let's format these names and concatenate with variable type.
unique(dat.final[, .(habitat_category, habitat_trait)])
dat.final[habitat_trait != "All", 
          habitat_trait := paste0(habitat_category, ".", 
                                  habitat_trait)]


```
The variable `habitat_trait` will allow for easy subgroup filtering as we fit our models.

## Set up model guide {#sec-model-guide}
Now let's create all the combinations of models we want to run. The function `data.table::CJ` is similar to `expand.grid` and makes all combinations of the input vectors. We'll do this in two steps to avoid having to filter out some combinations that we don't have sample size to run.

```{r}
#| eval: true
#| echo: true
#| output: true

# # Do the guide in 2 steps. We don't have enough data for class * habitat sub-analyses.
guide <- CJ(analysis_group = unique(dat.final$analysis_group),
            moderator = c("1", "log_mass"),
            class = c("Birds", "Mammals", "Reptiles", "All"),
            habitat_trait = "All",
            prey_range = c("All", "inside"),
            non_phylo_species = c("yes", "no"),
            phylo_species = c("yes", "no"),
            only_dominant_effect_size = c("yes", "no")
)
guide <- guide[!(moderator == "log_mass" & class != "All"), ]

guide2 <- CJ(analysis_group = unique(dat.final$analysis_group),
             moderator = "1",
             habitat_trait = c("locomotion_volant.non_volant", 
                               "locomotion_volant.volant", 
                               "ground_or_burrow_resting_or_nesting.other",
                               "ground_or_burrow_resting_or_nesting.ground", 
                               "Foraging_habitat_ground.other",
                               "Foraging_habitat_ground.ground", 
                               "continent_island.island", 
                               "continent_island.mainland"),
             class = c("All"),
             prey_range = c("All"),
             non_phylo_species = c("yes", "no"),
             phylo_species = c("yes", "no"),
             only_dominant_effect_size = c("no")
             )

guide <- rbind(guide, guide2)

```

Now we'll add a unique ID that we'll use to compare alternative random effects per model series (`model_comparison_id`). And we'll formulate strings which we'll evaluate as R code for our random effects (`random_effect`), to filter our dataset (`exclusion`) and for our fixed effect formula (`formula`):

```{r}
#| eval: true
#| echo: true
#| output: false

guide[, model_comparison_id := paste0("model_comp_", .GRP), 
      by = .(analysis_group, class, moderator, prey_range,
             habitat_trait, only_dominant_effect_size)]

guide[, random_effect := "list(~1|article_id/Effect_size_ID"]

guide[, random_effect := ifelse(non_phylo_species == "yes",
                                paste0(random_effect, ", ~1|scientificName"), 
                                paste0(random_effect))]
guide[, random_effect := ifelse(phylo_species == "yes",
                                paste0(random_effect, ", ~1|phylo_species)"),
                                paste0(random_effect, ")"))]
guide

# Formulate exclusion formula:
guide[, exclusion := paste0("analysis_group == '", analysis_group, "'",
                            " & habitat_trait == '", habitat_trait, "'")]
guide[class != "All", exclusion := paste0(exclusion, " & class %in% '", class, "'")]

guide[prey_range == "inside", 
      exclusion := paste0(exclusion, " & prey_range_primary == 'inside'")]
guide[prey_range == "All"]

#
guide.m1 <- merge(guide,
               unique(dat[, .(analysis_group, analysis_effect_size)]),
               by = 'analysis_group',
               all.x = T)
guide.m1[is.na(analysis_effect_size)]
guide.m1

#
guide.m1[only_dominant_effect_size == "yes", 
         exclusion := paste0(exclusion, " & dominant_effect_size == 'yes'")]

#
guide.m1[, formula := ifelse(moderator == "1",
                             "~ 1",
                             paste("~", moderator))]


guide.m1[, model_id := paste0("model_", seq(1:.N))]

```

The resulting guide will execute 1,232 models and looks like this: 

```{r}
#| eval: true
#| echo: true
#| output: true

head(guide.m1)
```
The columns `exclusion`, `random_effect`, and `formula` are text strings that we will evaluate to filter the dataset and run models. 


Now we'll filter out models for which we have insufficient sample size. We'll do this by evaluating the exclusion strings as a subsetting operation on the dataset. We'll be very generous with sample size inclusion criteria and will fit models if there are at least 2 observations for intercept-only models and 5 observations for models with a continuous moderator (body mass).

```{r}
#| eval: true
#| echo: true
#| output: false

Ns <- list()
sub.dat <- c()

for(i in 1:nrow(guide.m1)){
  sub.dat <- dat.final[eval(parse(text = guide.m1[i, ]$exclusion)), ] # Evaluate `esclusion` string
  
  Ns[[i]] <- sub.dat[, .(n_species = uniqueN(scientificName),
                    n_articles = uniqueN(Article),
                    n_obs = .N,
                    model_id = guide[i, ]$model_id,
                    analysis_group = guide[i, ]$analysis_group,
                    class = guide[i, ]$class)]
  Ns[[i]]$model_id <- guide.m1[i, ]$model_id
  
}
Ns <- rbindlist(Ns)

guide.m2 <- merge(guide.m1, 
               Ns[, .(n_species, n_articles, n_obs, model_id)], 
               by = "model_id")

# More than 2 observations:
guide.m2 <- guide.m2[n_obs > 2, ]

# and at least 5 for continuous
guide.m2 <- guide.m2[!(moderator == "log_mass" & n_articles < 5), ]
head(guide.m2)

```

If there is only 1 article per analysis group we will drop 'article_id` from the random effects:

```{r}
#| eval: true
#| echo: true
#| output: true

guide.m2[n_articles == 1, random_effect := gsub("article_id/", "", random_effect)]
guide.m2

```

To facilitate model fitting and tidying we'll also set up a prediction grid for extracting predictions for our only moderator, log mass.

```{r}
#| eval: true
#| echo: true
#| output: false

unique(guide.m2$moderator)
grids <- list()
grids[["log_mass"]] <- data.table(log_mass = seq(from = min(dat.final$log_mass, na.rm = T),
                                                 to = max(dat.final$log_mass, na.rm = T),
                                                 by = .1))

```

## Prepare phylogeny {#sec-phylogeny}

Now we'll prepare a phylogeny using the `rotl` package [@rotl-cit]. This will download a `phylo` object and we will merge in the tip ID to our dataset (`dat.final`)

```{r}
#| eval: true
#| echo: true
#| output: true

dat.final[, spp_name_corrected := scientificName]
dat.final[scientificName == "Pampusana erythroptera",
         spp_name_corrected := "Gallicolumba erythroptera"]
nms <- unique(dat.final$spp_name_corrected)

nms_res <- rotl::tnrs_match_names(nms)
#
tree <- tol_induced_subtree(ott_ids = nms_res$ott_id)

plot(tree)

setDT(nms_res)
nms_res[, label := paste0(gsub(" ", "_", unique_name),
                          "_ott", ott_id)]

nms_res[, search_string := str_to_sentence(search_string)]

dat.final.m <- merge(dat.final,
                 nms_res[, .(search_string, label)],
                 by.x = "spp_name_corrected",
                 by.y = "search_string",
                 all.x = T)
nrow(dat.final.m) == nrow(dat.final)
setnames(dat.final.m, "label", "phylo_species")

```

## Run models {#sec-run-models}
Finally, let's run the models. We'll save the raw model objects in a `list()` object named `models`, the predictions in a `list()` object named `predictions` and tidied tables of model coefficients in a `list()` called `tidy_models`. 

In total, we have 528 models to run. Many of these will not fit because they have tiny sample sizes and complex non-independence structures (with the phylogeny and species identity random effects especially). However, we'll give it a try. The model fitting is done by `metafor::rma.mv`. 

```{r}
#| eval: true
#| echo: true
#| output: false

# Clean up the objects in the environment:
guide <- copy(guide.m2)
dat <- copy(dat.final.m)
rm(guide.m2, guide.m1, dat.final.m)

models <- list()
predictions <- list()
tidy_models <- list()

for(i in 1:nrow(guide)){
  
  sub_guide <- guide[i, ]
  sub_guide
  dat.sub <- dat[eval(parse(text = sub_guide$exclusion))]
  dat.sub
  #
  models[[i]] <- try({
      if(sub_guide$phylo_species == "yes"){
        
         rma.mv(yi_analysis, 
                V = vi_analysis,
                mods = as.formula(sub_guide$formula),
                random = eval(parse(text = sub_guide$random_effect)),
                dfs = "contain",
                test = "t",
                method = "REML",
                R =  list(phylo_species = createPhyloCorr(dat.sub$phylo_species, tree)),
                data = dat.sub) 
        
      }else{
        rma.mv(yi_analysis, 
                V = vi_analysis,
                mods = as.formula(sub_guide$formula),
                random = eval(parse(text = sub_guide$random_effect)),
                dfs = "contain",
                test = "t",
                method = "REML",
                data = dat.sub) 
      }
  })

  if(inherits(models[[i]], "try-error")){
    models[[i]] <- as.character(models[[i]])
  }else{
    
    if(sub_guide$moderator == "1"){
      predictions[[i]] <- predict(models[[i]]) |>
        as.data.frame() |>
        bind_cols(sub_guide) |>
        rename(lower_ci = ci.lb,
               upper_ci = ci.ub,
               lower_pi = pi.lb,
               upper_pi = pi.ub) |>
        setDT()
    }else{
      predictions[[i]] <- rma_predictions(models[[i]], grids[[sub_guide$moderator]]) |>
        rename("term" = sub_guide$moderator) |>
        bind_cols(sub_guide) |>
        rename(lower_ci = ci.lb,
               upper_ci = ci.ub,
               lower_pi = pi.lb,
               upper_pi = pi.ub) |>
        setDT()
    }
    
    tidy_models[[i]] <- tidy(models[[i]]) |>
      mutate(lower_ci = models[[i]]$ci.lb, 
             upper_ci = models[[i]]$ci.ub,
             df = unique(models[[i]]$ddf)) |>
      mutate(overfit = ifelse(any(models[[i]]$sigma2 == 0), "yes", "no"),
             aic = AIC(models[[i]])) |>
      bind_cols(as.data.frame(i2_ml(models[[i]])) |> t() ) |>
      bind_cols(sub_guide) |>
      setDT()
    
    cat(i, "/", nrow(guide), "\r")
    
    names(predictions)[i] <- sub_guide$model_id
    names(tidy_models)[i] <- sub_guide$model_id
    # names(models[i]) <- tidy_models[i]$model_id
    
  } 
}

```

## Model selection {#sec-model-selection}
Now we'll filter out overfit models by removing any with a sigma equal to 0 (which has been designated with the column `overfit`). Then we'll select the best model per analysis group, moderator, subgroup, and sensitivity analysis category (specified with `model_comparison_id`). 

```{r}
#| eval: true
#| echo: true
#| output: false

names(models) <- guide$model_id

setdiff(names(tidy_models[[2]]), names(tidy_models[[1]]))
tidy_models <- rbindlist(tidy_models, fill = TRUE)
tidy_models[overfit == "yes", ]
tidy_models <- tidy_models[overfit != "yes", ]

length(predictions)
predictions <- rbindlist(predictions, fill = TRUE)
predictions
names(models)


# >>> Select best model ---------------------------------------------------
tidy_models[, min_aic := min(aic),
            by = .(model_comparison_id)]

tidy_models <- tidy_models[min_aic == aic, ]
tidy_models[overfit == "yes", ]

predictions <- predictions[model_id %in% tidy_models$model_id]
predictions

# Now let's save these objects and reload in the next section:
saveRDS(tidy_models, "builds/meta_analysis/models/tidy_models.Rds")
saveRDS(predictions, "builds/meta_analysis/models/predictions.Rds")
saveRDS(dat, "builds/meta_analysis/models/analysis_data.Rds")

```

See the [next section](3.3-plot-models.qmd) for meta-analysis figures
