---
title: "Publication bias tests"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "3-publication-bias.html"
    # embed-resources: true
    code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---

Given that some of our analysis contain mixed effect sizes we elected to only conduct time-lag publication bias tests. Other publication bias tests (e.g., small sample size) require uniform measurements of effective sample size or variance and are not suitable for original effect size types. 

Time lag publication bias tests asks if significant effects have declined (or increased) since initial discovery, which is common when strong initial findings spur the development of a field. We thus regress our effect sizes against publication year. We'll do a similar approach with a model guide as we did [above](#sec-run-models). Except we'll put all of the code into a single code block. This workflow is derived from a full R script in `scripts/2_analysis/3 publication bias.R` that includes additional checks and tests and is fully interactive if you clone the repo.

```{r}
#| eval: true
#| echo: true
#| output: false

# Clean environment:
rm(list = ls())

# Load packages
library("data.table")
library("metafor")
library("ggplot2")
library("ggstance")
library("orchaRd")
library("patchwork")
library("readxl")
library("stringr")
library("sf")
library("mapview")
library("dplyr")
library("rotl")
library("broom.mixed")
library("here")

rma_predictions <- function(m, 
                            newgrid,
                            has_intercept = T){
  
  
  if(!is.data.frame(newgrid)){errorCondition("ERROR newgrid must be a data frame")}
  #create the new model matrix. 
  
  if(!all(unlist(lapply(names(newgrid), # lapply through names of newgrid to check that they're in formula
                        function(x) grepl(pattern=x,
                                          x = as.character(m$formula.mods)[-1]))))){
    errorCondition("ERROR: variables in newgrid are not in model formula")
  }
  
  
  # Drop levels that might be missing from the model...
  cols <- names(newgrid)
  coef_nms <- names(coef(m))
  temp <- c()
  
  if(has_intercept == F){
    
    for(i in 1:length(cols)){
      if(class(unlist(newgrid[, cols[i], with = F])) %in% c("factor", "character")){
        temp <- paste0(names(newgrid[, cols[i], with = F]),
                       unlist(newgrid[, cols[i], with = F]))
        newgrid <- newgrid[temp %in% coef_nms, ]
      }
    }
    
  }
  
  newgrid
  
  # Create prediction matrix
  predgrid <- (model.matrix(m$formula.mods, data=newgrid))
  predgrid
  
  if(any(grepl("intercept", colnames(predgrid), 
               ignore.case = TRUE))){
    #if intercept is present, remove it?
    predgrid <- predgrid[, -1]
  }
  
  # predict onto the new model matrix
  pred.out <- as.data.frame(predict(m, newmods=predgrid))
  
  #attach predictions to variables for plotting
  final.pred <- cbind(newgrid, pred.out)
  
  return(final.pred)
}


# Load data ---------------------------------------------------------------

dat <- fread("builds/meta_analysis/analysis_ready_dataset.csv")

dat[, article_id := paste(word(Article, 1, sep = "[[:space:][:punct:]]"), .GRP), by = .(Article)]
unique(dat[, .(Article, article_id)])
nrow(dat[duplicated(article_id)])
nrow(dat[duplicated(Article)])
# Must be equal. Good

unique(dat[, .(analysis_group, analysis_effect_size)])


dat.long <- copy(dat)

# >>> Extract publicaiton years ------------------------------------------
library("readr")

# parse_number(x = "Gerber, G.P. and Iverson, J.B., 2000. Turks and Caicos iguana, Cyclura carinata carinata. West Indian Iguanas: Status Survey and Conservation Action Plan. IUCN the World Conservation Union, Gland, Switzerland, pp.15-18.")
str_extract("Gerber, G.P. and Iverson, J.B., 2000. Turks and Caicos iguana, Cyclura carinata carinata. West Indian Iguanas: Status Survey and Conservation Action Plan. IUCN the World Conservation Union, Gland, Switzerland, pp.15-18.", 
            "\\d+")

dat.long[, year := str_extract(Article,
                               "\\d+")]
unique(dat.long$year)
dat.long[year == "6", ]$Article
dat.long[year == "6", year := 1974]

dat.long[year == "13", ]$Article
dat.long[year == "13", year := 2010]

dat.long[year == "16", ]$Article
dat.long[year == "16", year := 2021]

dat.long[year == "35", ]$Article
dat.long[year == "35", year := 2010]

dat.long[year == "104", ]$Article
dat.long[year == "104", year := 2007]

dat.long[year == "12", ]$Article
dat.long[year == "12", year := 2009]

dat.long[, year := as.numeric(year)]

# Because of mixed effect size types, let's just do time-lag publication bias?

dat.long[, year_scaled := scale(year)]

# Set up model guide -----------------------------------------------------------------

unique(dat.long$analysis_group)


guide <- CJ(analysis_group = unique(dat.long$analysis_group),
            moderator = c("year_scaled")
)

guide[, random_effect := "list(~1|article_id/Effect_size_ID"]

guide[, exclusion := paste0("analysis_group == '", analysis_group, "'")]
# guide[class != "All", exclusion := paste(exclusion, "& class == '", class, "'")]

guide[, formula := paste("~", moderator)]

guide

guide[, model_id := paste(analysis_group, moderator)]

# Run models --------------------------------------------------------------

bias_models <- list()
bias_tidy <- list()
i <- 1

for(i in 1:nrow(guide)){
  sub.dat <- dat.long[eval(parse(text = guide$exclusion[i]))]
  
  bias_models[[i]] <- try({
                            rma.mv(yi_analysis,
                             V = vi_analysis,
                             method = "REML",
                             test = "t",
                             dfs = "contain",
                             mods = as.formula(guide$formula[i]),
                             random = ~1 | article_id/Effect_size_ID,
                             data = sub.dat)
  })
  if(!inherits(bias_models[[i]], "try-error")){
    bias_tidy[[i]] <- bias_models[[i]] |> 
      tidy() |>
      mutate(sigma = min(bias_models[[i]]$sigma2),
             df = paste(bias_models[[i]]$ddf, collapse = ", "),
             lower_ci = bias_models[[i]]$ci.lb,
             upper_ci = bias_models[[i]]$ci.ub) |>
      bind_cols(guide[i, ]) |>
      setDT()
  }

  
}

# names(bias_models) <- guide$model_id
bias_models

# >>> Interpret -----------------------------------------------------------
# Bias:
bias_tidy <- rbindlist(bias_tidy)
bias_tidy
```

Now let's see if any have a significant coefficient for year, and if so, what direction effect sizes have trended over time:

```{r}
#| eval: true
#| echo: true
#| output: true

bias_tidy[p.value < 0.05, ]

```
Looks like there are no significant coefficients. The p-values for year range from:

```{r}
#| eval: true
#| echo: true
#| output: true

round(range(bias_tidy[term == "year_scaled", ]$p.value), 2)
```

and the coefficients range from:

```{r}
#| eval: true
#| echo: true
#| output: true

round(range(bias_tidy[term == "year_scaled", ]$estimate), 2)
```

