---
title: "Appendix: effect size conversion formulas"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "3.4-effect-size-conversions.qmd"
    # embed-resources: true
    code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---
```{r}
#| eval: true
#| echo: false
#| output: false

# Clean environment:
rm(list = ls())

# Load packages
library("data.table")
library("metafor")
library("ggplot2")
library("ggstance")
library("orchaRd")
library("patchwork")
library("readxl")
library("stringr")
library("sf")
library("mapview")
library("dplyr")
library("rotl")
library("broom.mixed")
library("here")
```

These effect size formulas are digitized in the function `convert_effect_sizes` which has them scripted into R code. The formulas are in a side car csv file in the repo located at 'scripts/functions/conversion_formulas.csv'

```{r}
#| eval: true
#| echo: true
#| output: false

source("scripts/functions/convert_effect_sizes.R")

formulas <- fread("scripts/functions/conversion_formulas.csv")

```

We'll show the formulas below and demonstrate the use of the function. Note in all cases we first need to convert to and from Hedges' g and Cohen's d (the uncorrected estimator of SMD). Hedges' g is the effect size estimator used in our analyses and is returned by default from `escalc`. This conversion looks like this:

$$
\begin{aligned}
&J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1} \\
&yi_d = yi_g * J \\
\end{aligned}
$${#eq-1}

# SMD ↔ lnOR 

## SMD → lnOR {#sec-smd-lnOR}

Let $yi_{lnOR}$ and $yi_{d}$ be the point estimate for the lnOR and Cohen's d effect sizes and $vi_{lnOR}$ and $vi_{d}$ be their sampling variances. The resulting estimate of Cohen's d was then converted using @eq-1 to Hedges' g. Note that the estimator of sampling variance for SMD follows `escalc` and uses the sampling variance of Cohen's d instead of Hedges' g, thus no further conversion is necessary.


$$
\begin{aligned}
&yi_{lnOR} = yi_d * \frac{\pi}{\sqrt{3}} \\
&vi_{lnOR} = vi_d * \frac{\pi^2}{3} \\
\end{aligned}
$${#eq-2}

In `R`:

```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "SMD",
                     to = "lnOR",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)

```


Now let's look at the transcribed formula for the point estimate conversion. `yi` and `vi` refer to the original effect size and sampling variance estimates (in this case of lnOR) and `yi_trans` and `vi_trans` refer to the converted estimates (in this case of SMD), `n1` and `n2` = $n_{1, 2}$, and `d` = $yi_d$. Note that our formula for $J$ uses the gamma distribution instead of $J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1}$:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "SMD" & to_effect == "lnOR"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

## lnOR → SMD {#sec-lnOR-smd}
To convert from lnOR ($yi_{lnOR}$) to Cohen's $d$:

$$
\begin{aligned}
&yi_d = yi_{lnOR} * \frac{\sqrt{3}}{\pi} \\
&vi_d = vi_{lnOR} * \frac{3}{\pi^2} \\
\end{aligned}
$${#eq-3}

In `R`:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "lnOR",
                     to = "SMD",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)
```

And the transcribed formulas in `R`:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "lnOR" & to_effect == "SMD"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

The variable names are the same as [above](#sec-smd-lnOR).

# SMD ↔ Zr
## SMD → Zr {#sec-smd-zr}

To convert from SMD to Zr, we first convert $d$ to the correlation coefficient $r$ and then to Zr, using a correction factor $a$ for cases where $n_1 \neq n_2$. Let $yi_{Zr}$ and $yi_d$ be the point estimate for Zr and Cohen's d:

$$
\begin{aligned}
&a = \frac{(n_1 + n_2)^2}{n_1*n_2} \\
&r = \frac{yi_d}{\sqrt{yi_d^2 + a}} \\
&r \leftarrow \min(1, \max(-1, r_i))\\
&yi_{Zr} = atanh(r)\\
\end{aligned}
$${#eq-4}

And the sampling variance of Zr ($vi_{Zr}$):
$$
\begin{aligned}
&n = n1 + n2 \\
&vi_{Zr} = \frac{1}{(n - 3)} \\
\end{aligned}
$${#eq-5}

In R:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "SMD",
                     to = "Zr",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)
```

And let's look at the `R` formulas. `yi` and `vi` refer to the original effect size and sampling variance estimates (in this case of SMD) and `yi_trans` and `vi_trans` refer to the converted estimates (in this case of Zr), `n1` and `n2` = $n_{1, 2}$, `a` = the correction factor $a$, and `d` = $yi_d$. Note that our formula for $J$ uses the gamma distribution instead of $J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1}$:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "SMD" & to_effect == "Zr"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

## Zr → SMD {#sec-zr-smd}
First, convert the point estimate of Zr ($yi_{Zr}$) to the correlation coefficient $r$ and calculate the variance of $r$ ($vi_r$:
$$
\begin{aligned}
&r = tanh(yi_{Zr}) \\
&vi_r = \frac{(1-r^2)^2}{(n-1)}
\end{aligned}
$${#eq-6}

Then calculate the point estimate ($yi_d$) and variance ($vi_d$) of Cohen's d:

$$
\begin{aligned}
&yi_d = \frac{2*r}{\sqrt{1-r^2}} \\
&vi_d = \frac{4*vi_r}{(1-r^2)^3} \\
\end{aligned}
$${#eq-7}

In R:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "Zr",
                     to = "SMD",
                     yi = 0.9,
                     vi = 0.01,
                     n = 50)

```

And here are the transcribed formulas used:
```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "Zr" & to_effect == "SMD"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])

```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```
The variable names are the same as [above](#sec-smd-zr).