---
title: "Appendix: effect size conversion formulas"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "3.4-effect-size-conversions.qmd"
    # embed-resources: true
    code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---
```{r}
#| eval: true
#| echo: false
#| output: false

# Clean environment:
rm(list = ls())

# Load packages
library("data.table")
library("metafor")
library("ggplot2")
library("ggstance")
library("orchaRd")
library("patchwork")
library("readxl")
library("stringr")
library("sf")
library("mapview")
library("dplyr")
library("rotl")
library("broom.mixed")
library("here")
```

We converted the effect sizes into common types based on the formulas in [@borenstein2021], which we digitized in our function `convert_effect_sizes`. The formulas are in an executable sidecar csv file in the repo located at 'scripts/functions/conversion_formulas.csv'

```{r}
#| eval: true
#| echo: true
#| output: false

source("scripts/functions/convert_effect_sizes.R")

formulas <- fread("scripts/functions/conversion_formulas.csv")

```

We'll show the formulas below, demonstrate the use of the function, and the digitized `R` formulas. 

:::callout-important
Note that in all cases, the conversion formulas below are based on conversions from Cohen's d (the uncorrected estimator of SMD). As such, SMD estimates had to first be transformed to Cohen's d, and the outputs of conversions had to be transformed to Hedges' g. Converting from Hedges' g (the default estimator returned from `escalc`) is done as follows, with $n_1$, $n_2$ equivalent to the sample sizes of groups 1 and 2, $yi_g$ and $yi_d$ the effect size point estimate sfor Hedges' g and Cohen's d, and $J$ a small-sample bias correction factor. 

$$
\begin{aligned}
&J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1} \\
\end{aligned}
$${#eq-1}

$$
\begin{aligned}
&yi_d = \frac{yi_g}{J} \\
\end{aligned}
$${#eq-2}

$$
\begin{aligned}
&yi_g = J * yi_d \\
\end{aligned}
$${#eq-3}

Note that the estimator of sampling variance for SMD, following `escalc`, is Cohen's d's sampling variance estimator and thus did not need to be converted with $J$.
:::

# SMD ↔ lnOR 

## SMD → lnOR {#sec-smd-lnOR}

Let $yi_{lnOR}$ and $yi_{d}$ be the point estimate for the lnOR and Cohen's d effect sizes and $vi_{lnOR}$ and $vi_{d}$ be their sampling variances. The resulting estimate of Cohen's d was then converted using @eq-3 to Hedges' g. 


$$
\begin{aligned}
&yi_{lnOR} = yi_d * \frac{\pi}{\sqrt{3}} \\
&vi_{lnOR} = vi_d * \frac{\pi^2}{3} \\
\end{aligned}
$${#eq-4}

In `R`:

```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "SMD",
                     to = "lnOR",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)

```


Now let's look at the transcribed formula for the point estimate conversion. `yi` and `vi` refer to the original effect size and sampling variance estimates (in this case of lnOR) and `yi_trans` and `vi_trans` refer to the converted estimates (in this case of SMD), `n1` and `n2` = $n_{1, 2}$, and `d` = $yi_d$. Note that our formula for $J$ uses the gamma distribution instead of $J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1}$:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "SMD" & to_effect == "lnOR"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

## lnOR → SMD {#sec-lnOR-smd}
To convert from lnOR ($yi_{lnOR}$) to Cohen's $d$:

$$
\begin{aligned}
&yi_d = yi_{lnOR} * \frac{\sqrt{3}}{\pi} \\
&vi_d = vi_{lnOR} * \frac{3}{\pi^2} \\
\end{aligned}
$${#eq-5}

$yi_d$ was then converted to Hedges' g using @eq-3.

In `R`:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "lnOR",
                     to = "SMD",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)
```

And the transcribed formulas in `R`:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "lnOR" & to_effect == "SMD"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

The variable names are the same as [above](#sec-smd-lnOR).

# SMD ↔ Zr
## SMD → Zr {#sec-smd-zr}

To convert from SMD to Zr, we used @eq-2 to convert Hedges' g to Cohen's d. We then converted $yi_d$ to the correlation coefficient $r$ and then to Zr, using a correction factor $a$, to account for cases where $n_1 \neq n_2$. Let $yi_{Zr}$ and $yi_d$ be the point estimate for Zr and Cohen's d:

$$
\begin{aligned}
&a = \frac{(n_1 + n_2)^2}{n_1*n_2} \\
&r = \frac{yi_d}{\sqrt{yi_d^2 + a}} \\
&r \leftarrow \min(1, \max(-1, r_i))\\
&yi_{Zr} = atanh(r)\\
\end{aligned}
$${#eq-6}

And the sampling variance of Zr ($vi_{Zr}$):
$$
\begin{aligned}
&n = n1 + n2 \\
&vi_{Zr} = \frac{1}{(n - 3)} \\
\end{aligned}
$${#eq-7}

In R:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "SMD",
                     to = "Zr",
                     yi = 0.9,
                     vi = 0.01,
                     n1 = 15,
                     n2 = 15)
```

And let's look at the `R` formulas. `yi` and `vi` refer to the original effect size and sampling variance estimates (in this case of SMD) and `yi_trans` and `vi_trans` refer to the converted estimates (in this case of Zr), `n1` and `n2` = $n_{1, 2}$, `a` = the correction factor $a$, and `d` = $yi_d$. Note that our formula for $J$ uses the gamma distribution instead of $J = 1 - \frac{3}{4 * (n_1 + n_2 - 2) - 1}$:

```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "SMD" & to_effect == "Zr"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])
```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```

## Zr → SMD {#sec-zr-smd}
First, convert the point estimate of Zr ($yi_{Zr}$) to the correlation coefficient $r$ and calculate the variance of $r$ ($vi_r$):
$$
\begin{aligned}
&r = tanh(yi_{Zr}) \\
&vi_r = \frac{(1-r^2)^2}{(n-1)}
\end{aligned}
$${#eq-8}

Then calculate the point estimate ($yi_d$) and variance ($vi_d$) of Cohen's d:

$$
\begin{aligned}
&yi_d = \frac{2*r}{\sqrt{1-r^2}} \\
&vi_d = \frac{4*vi_r}{(1-r^2)^3} \\
\end{aligned}
$${#eq-9}

Cohen's d ($yi_d$) was then converted to Hedges' g following @eq-3.

In R:
```{r}
#| eval: true
#| echo: true
#| out-width: "70%"
convert_effect_sizes(from = "Zr",
                     to = "SMD",
                     yi = 0.9,
                     vi = 0.01,
                     n = 50)

```

And here are the transcribed formulas used:
```{r}
#| echo: false
#| eval: true

sub_f <- formulas[from_effect == "Zr" & to_effect == "SMD"]$formula
sub_f <- strsplit(sub_f, split = "; ")

print(sub_f[[1]])

```

And for sampling variance:

```{r}
#| echo: false
#| eval: true

print(sub_f[[2]])
```
The variable names are the same as [above](#sec-smd-zr).