---
title: "Where does the evidence come from?"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "4-map.html"
    # embed-resources: true
    # code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---

Let's see you where the evidence collected about cats comes from. The code block is below but we've hidden it by default to simplify readability. Click the `Code` button if you want to take a look.  

```{r}
#| eval: true
#| echo: true
#| output: false
#| code-fold: true

rm(list = ls())

library("data.table")
library("sf")
library("mapview")
library("ggplot2")
library("stringr")
library("dplyr")

# Load the datasets this needs:
sys_rev <- fread("builds/systematic_review/systematic_review_tidy.csv")
dat <- fread("builds/meta_analysis/analysis_ready_dataset.csv")

# sys_rev$Study_lat
sys_rev[Article_simple %in% dat$Article]

setdiff(dat$Article, sys_rev$Article_simple)

dat$Article %in% sys_rev[grepl("Population", evidence_type), ]$Article_simple

sys_rev[grepl("Population", evidence_type), 
        in_meta := ifelse(Article_simple %in% dat$Article, "yes", "no")]

# Fix some text parsing issues manually:
sys_rev[grepl("Does aerial baiting for controlling feral cats in a heterogeneous landscape confer benefits",
              Article_simple) &
          grepl("Population", evidence_type), 
        in_meta := "yes"]

sys_rev[grepl("Are dingoes a trophic regulator in arid Australia?", 
              Article_simple, ignore.case = T),]
sys_rev[grepl("Population", evidence_type) &
          grepl("Are dingoes a trophic regulator in arid Australia?",
              Article_simple, ignore.case = T), 
        in_meta := "yes"]

#
sys_rev[grepl("Does a top‐predator provide an endangered rodent with refuge from an invasive mesopredator?", 
              Article_simple, ignore.case = T),]
sys_rev[grepl("Does a top‐predator provide an endangered rodent with refuge from an invasive mesopredator?",
              Article_simple, ignore.case = T)  &
          grepl("Population", evidence_type), 
        in_meta := "yes"]


sys_rev[is.na(as.numeric(Study_lat))]
sys_rev[, Study_lat := gsub("−", "-", Study_lat)]
sys_rev[, Study_long := gsub("−", "-", Study_long)]
sys_rev[is.na(as.numeric(Study_lat))]
sys_rev[is.na(as.numeric(Study_long))]$Study_long
sys_rev[, Study_long := str_trim(Study_long)]
sys_rev[, Study_long := gsub("﻿", "", Study_long)]
sys_rev[is.na(as.numeric(Study_long))]$Study_long

sys_rev[, `:=` (Study_lat = as.numeric(Study_lat),
                Study_long = as.numeric(Study_long))]
sys_rev

# >>> Plot ----------------------------------------------------

unique(sys_rev$Evidence_category)
sys_rev[Evidence_category == "Control program", Evidence_category := "Population"]
#
sys_rev
sys_rev[, synth_fill := fcase(grepl("Population", evidence_type) & has_data == "yes", paste(evidence_type, "with data"),
                                grepl("Population", evidence_type) & has_data == "no", paste(evidence_type, "without data"))]

sys_rev[is.na(synth_fill), synth_fill := evidence_type]
sys_rev[of_quality == "yes",]
sys_rev[of_quality == "yes", synth_fill := paste(synth_fill, "and qualities")]

unique(sys_rev$synth_fill)

fill_pal <- c(`Predation in support` = "#a6d3a0", `Predation not in support` = "#40531b", 
              `Population not in support without data` = "indianred4", `Population not in support with data` = "indianred", 
              `Population not in support with data and qualities` = "indianred", 
              `Population in support without data` = "dodgerblue4", `Population in support with data` = "dodgerblue",
              `Population in support with data and qualities` = "dodgerblue"
)

unique(sys_rev$in_meta)
unique(sys_rev[is.na(in_meta)]$synth_fill)
sys_rev[is.na(in_meta), in_meta := "no"]
sys_rev$in_meta <- factor(sys_rev$in_meta,
                          levels = c('no', 'yes'))

sys_rev[, in_meta_num := as.numeric(in_meta)]
sys_rev[, in_meta_num := fcase(in_meta_num == 1, 2,
                               in_meta_num == 2, 3)]

unique(sys_rev[in_meta_num == 3, ]$synth_fill)

sys_rev.spat <- sys_rev %>%
  filter(!is.na(Study_long)) %>%
  st_as_sf(coords = c("Study_long", "Study_lat"),
           crs = 4326)

sys_rev.spat$synth_fill <- factor(sys_rev.spat$synth_fill,
                                  levels = names(fill_pal))

sys_rev.spat <- sys_rev.spat %>%
  mutate(of_quality_color = ifelse(of_quality == "yes", "gold", "black")) 


```

Otherwise, here are interactive maps of all locations where quantitative evidence was found. 

```{r}
#| eval: true
#| echo: true
#| output: true
#| code-fold: true

m <- sys_rev.spat %>% 
  select(scientificName, class, Article_simple, 
         Study_location, synth_fill,
         of_quality, in_meta, in_meta_num) %>%
  rename(`In meta-analysis` = in_meta,
         `Has qualities` = of_quality,
         `Evidence type` = synth_fill,
         `Species` = scientificName,
         `Article` = Article_simple,
         Location = Study_location,
         `point size` = in_meta_num) %>%
  mapview(.,
        zcol = "Evidence type",
        col.regions = (fill_pal),
        color = sys_rev.spat$of_quality_color,
        cex = "point size",
        layer.name = "Evidence type") 

```

**For all species:**

```{r}
#| eval: true
#| echo: false
#| output: true
#| code-fold: false
#| label: Figure 1
#| fig-cap: "Quantitative studies linking cats to the endangerment of vertebrate species. Larger points indicate studies that could be included in meta-analysis. Gold borders indicate studies with qualities."

m

```


```{r}
#| eval: true
#| echo: true
#| output: true
#| code-fold: true


# PREPARE MAMMALS

sub_dat <- sys_rev.spat %>% 
  filter(class == "Mammals")

sub_fill <- fill_pal[names(fill_pal) %in% droplevels(sub_dat$synth_fill)]

m <- sub_dat %>%
  select(scientificName, class, Article_simple, 
         Study_location, synth_fill,
         of_quality, in_meta, in_meta_num) %>%
  rename(`In meta-analysis` = in_meta,
         `Has qualities` = of_quality,
         `Evidence type` = synth_fill,
         `Species` = scientificName,
         `Article` = Article_simple,
         Location = Study_location,
         `point size` = in_meta_num) %>%
  mapview(zcol = "Evidence type",
        color = sub_dat$of_quality_color,
        col.regions = (fill_pal),
        cex = "point size",
        layer.name = "Evidence type")

```

Here is where all evidence concerning **mammal prey** came from:
```{r}
#| eval: true
#| echo: false
#| output: true
#| code-fold: false
#| label: Figure 2
#| fig-cap: "Quantitative studies linking cats to the endangerment of mammal species. Larger points indicate studies that could be included in meta-analysis. Gold borders indicate studies with qualities."
m

```


```{r}
#| eval: true
#| echo: true
#| output: true
#| code-fold: true
# PREPARE BIRDS
sub_dat <- sys_rev.spat %>% 
  filter(class == "Birds")

sub_fill <- fill_pal[names(fill_pal) %in% droplevels(sub_dat$synth_fill)]

m <- sub_dat %>%
  select(scientificName, class, Article_simple, 
         Study_location, synth_fill,
         of_quality, in_meta, in_meta_num) %>%
  rename(`In meta-analysis` = in_meta,
         `Has qualities` = of_quality,
         `Evidence type` = synth_fill,
         `Species` = scientificName,
         `Article` = Article_simple,
         Location = Study_location,
         `point size` = in_meta_num) %>%
  mapview(zcol = "Evidence type",
        color = sub_dat$of_quality_color,
        col.regions = (fill_pal),
        cex = "point size",
        layer.name = "Evidence type")

```

And here are **birds:**

```{r}
#| eval: true
#| echo: false
#| output: true
#| code-fold: false
#| label: Figure 3
#| fig-cap: "Quantitative studies linking cats to the endangerment of bird species. Larger points indicate studies that could be included in meta-analysis. Gold borders indicate studies with qualities."
m
```


```{r}
#| eval: true
#| echo: true
#| output: true
#| code-fold: true
# REPTILES
sub_dat <- sys_rev.spat %>% 
  filter(class == "Reptiles")

sub_fill <- fill_pal[names(fill_pal) %in% droplevels(sub_dat$synth_fill)]

m <- sub_dat %>%
  select(scientificName, class, Article_simple, 
         Study_location, synth_fill,
         of_quality, in_meta, in_meta_num) %>%
  rename(`In meta-analysis` = in_meta,
         `Has qualities` = of_quality,
         `Evidence type` = synth_fill,
         `Species` = scientificName,
         `Article` = Article_simple,
         Location = Study_location,
         `point size` = in_meta_num) %>%
  mapview(zcol = "Evidence type",
        color = sub_dat$of_quality_color,
        col.regions = (fill_pal),
        cex = "point size",
        layer.name = "Evidence type")

```

**Reptiles:**
```{r}
#| eval: true
#| echo: false
#| output: true
#| code-fold: false
#| label: Figure 4
#| fig-cap: "Quantitative studies linking cats to the endangerment of reptile species. Larger points indicate studies that could be included in meta-analysis. Gold borders indicate studies with qualities."
m
```

```{r}
#| eval: true
#| echo: true
#| output: true
#| code-fold: true
# AMPHIBIANS
sub_dat <- sys_rev.spat %>% 
  filter(class == "Amphibians")

sub_fill <- fill_pal[names(fill_pal) %in% droplevels(sub_dat$synth_fill)]

m <- sub_dat %>%
  select(scientificName, class, Article_simple, 
         Study_location, synth_fill,
         of_quality, in_meta, in_meta_num) %>%
  rename(`In meta-analysis` = in_meta,
         `Has qualities` = of_quality,
         `Evidence type` = synth_fill,
         `Species` = scientificName,
         `Article` = Article_simple,
         Location = Study_location,
         `point size` = in_meta_num) %>%
  mapview(zcol = "Evidence type",
        color = sub_dat$of_quality_color,
        col.regions = (fill_pal),
        cex = "point size",
        layer.name = "Evidence type")

```
And, last but not least, **Amphibians:**

```{r}
#| eval: true
#| echo: false
#| output: true
#| code-fold: false
#| label: Figure 5
#| fig-cap: "Quantitative studies linking cats to the endangerment of amphibian species. Larger points indicate studies that could be included in meta-analysis. Gold borders indicate studies with qualities."
m
```

