---
title: "Timeline analysis"
date: today
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 6
    toc-float: true
    toc-expand: true
    toc-title: "**CONTENTS**"
    output-file: "5-timeline-analysis.qmd"
    # embed-resources: true
    code-fold: show
    code-link: true
    code-tools: true
    # number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "10"
    mainfont: "Helvetica"
    monofont: "Fira Code"
    # fontcolor: ""
    page-layout: article
    code-overflow: wrap
    df_print: paged
    theme: cosmo
    code-line-numbers: false
    grid:
      margin-width: 250px
crossref:
  fig-title: Figure     # (default is "Figure")
  tbl-title: Table     # (default is "Table")
  title-delim: "—"     # (default is ":")
  fig-prefix: Fig.     # (default is "Figure")
  tbl-prefix: Table.   # (default is "Table")
editor_options: 
  chunk_output_type: console
execute:
  warning: false
  message: false
  tidy: true
bibliography: references.bib
---

Now let's look at the timeline of the last records of extinct populations attributed to cat predation and the estimated arrival times of cats at those locations. These are reported based on the following categories based on population's estimated extinction year and estimated cat arrival year:

- Extinction before the arrival of cats
- Extinction overlaps arrival of cats
- Extinction within 0-100 years of cat arrival
- Extinction 101-200 years after cat arrival
- Extinction >200 years after cat arrival

This workflow corresponds to the full script at `scripts/2_analysis/5 timeline analysis.R`. Let's start by cleaning the environment and loading our data:

```{r}
#| eval: true
#| echo: true
#| output: true

rm(list = ls())

library("readxl")
library("data.table")
library("sf")
library("mapview")
library("rnaturalearth")
library("ggplot2")
library("dplyr")
library("plotly")
library("writexl")

theme_lundy <- theme_bw()+
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        strip.background = element_blank())

# Load data 
data_evidence <- read_xlsx("data/timeline/Cat timeline.xlsx") |> setDT()

```

Now let's do some data cleaning and create the categories listed above.

```{r}
#| eval: true
#| echo: true
#| output: true

#For now exclude one row of Conilurus capricornensis that has NA values
data_evidence <- data_evidence[!is.na(Prey_extirpation_year_earliest), ]

#Exclude Fregilupus varius (the IUCN does not claim that cats caused the extinction of this species)
data_evidence <- data_evidence[scientificName != "Fregilupus varius", ]

#Remove all “uncertain” rows
data_evidence <- data_evidence[is.na(AU_ONLY_Record_certainty) | 
                                 AU_ONLY_Record_certainty == "Confirmed"]

#How many species are in the dataset?

#Change status category to IUCN threat status
iucn_status <- read.csv("data/timeline/iucn_status.csv")
setDT(iucn_status)

full_data <- merge(iucn_status, data_evidence, by.x = "Scientific.name", by.y = "scientificName")

setnames(full_data, "Scientific.name", "Scientific name")
setnames(full_data, "Redlist.category", "Redlist category")

full_data[Class == "AVES", Class :="Birds"]
full_data[Class == "MAMMALIA", Class :="Mammals"]
full_data[Class == "REPTILIA", Class :="Reptiles"]

full_data[, IUCN_status:= NA_character_ ]
full_data[`Redlist category` == "Least Concern", IUCN_status:= "(LC)"]
full_data[`Redlist category` == "Near Threatened", IUCN_status:= "(NT)"]
full_data[`Redlist category` == "Vulnerable", IUCN_status:= "(VU)"]
full_data[`Redlist category` == "Endangered", IUCN_status:= "(EN)"]
full_data[`Redlist category` == "Critically Endangered", IUCN_status:= "(CR)"]
full_data[`Redlist category` == "Extinct in the Wild", IUCN_status:= "(EW)"]
full_data[`Redlist category` == "Extinct", IUCN_status:= "(EX)"]

#Assign color palette for species status
#IUCN palette
IUCN_palette <- list (
  "(LC)" = "#60c759",
  "(NT)" = "#A2CD5A",#"#cde321"
  "(VU)" = "#FFD700", #"#f9e90c",
  "(EN)" = "#fc803d",
  "(CR)" = "#d91801",
  "(EW)" = "#552243",
  "(EX)" = "#000000"
)

full_data <- full_data[, .(Class, Order, Family, Genus, `Scientific name`, `Redlist category`, IUCN_status,
                           Extinction_extirpation, Prey_extirpation_year_earliest, Prey_extirpation_year_earliest_CE_BP,
                           Prey_extirpation_year_latest, Prey_extirpation_year_latest_CE_BP,
                           Prey_extinction_timing_reference, Prey_quote, 
                           Cat_arrival_year_earliest, Cat_arrival_year_earliest_CE_BP, 
                           Cat_arrival_year_latest, Cat_arrival_year_latest_CE_BP, 
                           Cat_arrival_timing_reference, Cat_quote,
                           Location, Country, Latitude, Longitude, Realm, Systems)]

#Remove all LC species
full_data <- full_data[`Redlist category` != "Least Concern",]
full_data[, .N, by=`Redlist category`]
full_data[, .N, by= IUCN_status]

#How many are EX?
spp_list <- unique(full_data[`Redlist category`== "Extinct" | `Redlist category` == "Extinct in the Wild",
                             .(`Scientific name`, `Redlist category`)]) #There are 84 species

#Separate between extinct (EX + EW) and extant threatened species
full_data[, status:= NA_character_]
full_data[, status:= ifelse(`Redlist category`== "Extinct" | `Redlist category`== "Extinct in the Wild", "Extinct", "Extant")]
length(unique(full_data[status=="Extinct", `Scientific name`])) #Okay

# 3. Convert dates from CE to BP--------------------------------------------------
#Earliest prey extirpation → BP (min prey extinction)
full_data[, min_prey_extirpation_year_BP :=
            ifelse(Prey_extirpation_year_earliest_CE_BP == "CE",
                   1950 - Prey_extirpation_year_earliest,
                   Prey_extirpation_year_earliest)
]

#Latest prey extirpation → BP (max prey extinction)
full_data[, max_prey_extirpation_year_BP :=
            ifelse(Prey_extirpation_year_latest_CE_BP == "CE",
                   1950 - Prey_extirpation_year_latest,
                   Prey_extirpation_year_latest)
]

#Earliest cat arrival → BP (min cat arrival)
full_data[, min_cat_arrival_year_BP :=
            ifelse(Cat_arrival_year_earliest_CE_BP == "CE",
                   1950 - Cat_arrival_year_earliest,
                   Cat_arrival_year_earliest)
]

#Latest cat arrival → BP (max cat arrival)
full_data[, max_cat_arrival_year_BP :=
            ifelse(Cat_arrival_year_latest_CE_BP == "CE",
                   1950 - Cat_arrival_year_latest,
                   Cat_arrival_year_latest)
]

#Define time differences (prey extinction − cat arrival)
full_data[, `:=`(
  min_difference = min_prey_extirpation_year_BP - max_cat_arrival_year_BP,
  max_difference = max_prey_extirpation_year_BP - min_cat_arrival_year_BP
)]

full_data[, median_difference := (min_difference + max_difference) / 2]

#Flip direction
full_data[, `:=`(
  min_difference = -min_difference,
  max_difference = -max_difference,
  median_difference = -median_difference
)]

full_data[, prey_window :=
            min_prey_extirpation_year_BP - max_prey_extirpation_year_BP]

full_data[, cat_window :=
            min_cat_arrival_year_BP - max_cat_arrival_year_BP]

full_data[, cat_arrival := NA_character_]

# Before arrival
full_data[min_difference < 0 & max_difference < 0,
          cat_arrival := "Extinction before arrival"]

# Overlaps arrival
full_data[min_difference < 0 & max_difference >= 0,
          cat_arrival := "Extinction overlaps arrival"]

# 0–100 years
full_data[min_difference >= 0 & min_difference <= 100,
          cat_arrival := "Extinction within 100 years of arrival"]

# 101-200 years
full_data[min_difference > 100 & min_difference <= 200,
          cat_arrival := "Extinction 101–200 years after arrival"]

# >200 years
full_data[min_difference > 200,
          cat_arrival := "Extinction 201 years or more after arrival"]

full_data$cat_arrival <- factor(full_data$cat_arrival,
                                levels = c(
                                  "Extinction 201 years or more after arrival",
                                  "Extinction 101–200 years after arrival",
                                  "Extinction within 100 years of arrival",
                                  "Extinction overlaps arrival",
                                  "Extinction before arrival"
                                ))

full_data <- full_data[, .(Class, Order, Family, Genus, `Scientific name`, status, Extinction_extirpation,
                           `Redlist category`, IUCN_status, Location, Country, Latitude, Longitude, Realm, Systems,
                           Prey_extirpation_year_earliest, Prey_extirpation_year_earliest_CE_BP,
                           Prey_extirpation_year_latest, Prey_extirpation_year_latest_CE_BP,
                           Prey_extinction_timing_reference, Prey_quote, 
                           Cat_arrival_year_earliest, Cat_arrival_year_earliest_CE_BP, 
                           Cat_arrival_year_latest, Cat_arrival_year_latest_CE_BP, 
                           Cat_arrival_timing_reference, Cat_quote,
                           min_prey_extirpation_year_BP, max_prey_extirpation_year_BP, 
                           min_cat_arrival_year_BP, max_cat_arrival_year_BP, min_difference, max_difference,
                           median_difference, prey_window, cat_window, cat_arrival
)]

full_data <- full_data[order(full_data$Class), ]
table(full_data$cat_arrival)
```
That is the number of extinct populations per cat arrival category. 

Now let's summarize this dataset and create the main text figure:

```{r}
#| eval: true
#| echo: true
#| output: true

# 4. Summaries and figures--------------------------------------------------
spp_sum_ex <- unique(full_data[status == "Extinct", 
                               .(Class, `Scientific name`, 
                                 `Redlist category`, status, 
                                 cat_arrival, Country)])


ex_records <- full_data[status == "Extinct", ]

#Some summaries
sum_ex <- full_data[status=="Extinct", .N, by= .(Class, cat_arrival)]
sum_ex[, perc:= N / sum(N) * 100]
sum_ex[, sum(perc), by = cat_arrival]
sum_ex[, perc_class:= N / sum(N) * 100, by = Class]

setnames(sum_ex, "cat_arrival", "Cat arrival")
setnames(sum_ex, "N", "Populations")
setnames(sum_ex, "perc", "Total contribution (%)")
setnames(sum_ex, "perc_class", "Contribution by class (%)")

arrival_levels <- c(
  "Extinction before arrival",
  "Extinction overlaps arrival",
  "Extinction within 100 years of arrival",
  "Extinction 101–200 years after arrival",
  "Extinction 201 years or more after arrival"
)

sum_ex[, `Cat arrival` := factor(`Cat arrival`,
                                 levels = arrival_levels)]

setorder(sum_ex, `Cat arrival`, `Cat arrival`)
setorder(sum_ex, Class, -Class)

#Save dataset
writexl::write_xlsx(sum_ex, "builds/timeline/cat_extinctions_summary.xlsx")

#Let's make some summary figures
sum_ex[, `Cat arrival` := factor(`Cat arrival`,
                                 levels = arrival_levels)]

#Set color palette
plot_col<-c(
  "Extinction 201 years or more after arrival" = "#a2a2a2",
  "Extinction 101–200 years after arrival" = "#828282",
  "Extinction within 100 years of arrival" = "dodgerblue4",
  "Extinction overlaps arrival" = "grey20",
  "Extinction before arrival" = "black"
)

#Barplot by class & extinctions only
ggplot(sum_ex, 
       aes(x= Class, y=Populations, fill= `Cat arrival`))+
  geom_col(width = 0.8, position = position_dodge(width = 0.9))+
  scale_fill_manual(values = plot_col)+
  labs(x="", y="Number of attributed extinctions", fill= "")+
  theme_lundy +
  theme(legend.position = "inside",      
        legend.position.inside = c(0.8, 0.8))
```

Now let's look at all individual species:

```{r}
#| eval: true
#| echo: true
#| output: true
#| fig-width: 12
#| fig-height: 8

#Forest plot - species extinction and cat arrival times
plot_dt <- copy(full_data[`Redlist category`== "Extinct" | `Redlist category` == "Extinct in the Wild", ])

#Add a * to point out EW species
plot_dt[, sci_label:= ifelse(IUCN_status == "(EW)", paste(`Scientific name`, "*"), `Scientific name`)]

# Order species by Class, then alphabetically
plot_dt[, sci_label :=
          factor(sci_label,
                 levels = plot_dt[
                   order(Class, sci_label),
                   unique(sci_label)
                 ])
]

# Numeric index for background stripes
plot_dt[, x_id := as.numeric(sci_label)]

class_bounds <- plot_dt[, .(
  xmin = min(x_id) - 0.5,
  xmax = max(x_id) + 0.5
), by = Class]

class_labels <- plot_dt[, .(
  x = mean(x_id)
), by = Class]

#These three species don't show in the forest plot because there is no uncertainty in the cat and prey window. The max and min diff are the same, therefore the segment starts and end in the same place and becomes 0.
#To fix it add in the plot a tiny segment for these cases.
issue <- plot_dt[`Scientific name` %in% c("Todiramphus cinnamominus", "Cryptoblepharus egeriae", "Lepidodactylus listeri"), ]

ggplot() +
  # stripes
  geom_rect(
    data = plot_dt[, .(
      xmin = x_id - 0.5,
      xmax = x_id + 0.5,
      stripe = x_id %% 2)],
    aes(xmin = xmin, xmax = xmax,
        ymin = -Inf, ymax = Inf,
        fill = factor(stripe)),
    alpha = 0.4) +
  scale_fill_manual(values = c("white", "grey90"), guide = "none") +
  # class separators
  geom_vline(
    data = class_bounds,
    aes(xintercept = xmax),
    linetype = "dashed",
    colour = "grey30") +
  # uncertainty bars
  geom_segment(
    data = plot_dt,
    aes(x = sci_label,
        xend = sci_label,
        y = min_difference,
        yend = max_difference),
    linewidth = 1) +
  geom_hline(yintercept = 0) +
  geom_text(
    data = class_labels,
    aes(x = x, y = max(plot_dt$max_difference, na.rm = TRUE) + 40,
        label = Class),
    fontface = "bold",
    size = 8)+
  geom_segment(
    data = plot_dt[min_difference == max_difference],
    aes(x = sci_label, xend = sci_label,
        y = min_difference - 1, yend = max_difference + 1),
    linewidth = 1) +
  labs(
    y = "Years between cat arrival and last record",
    x = NULL) +
  theme_lundy +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, face = "italic"),
    axis.text = element_text(size=12),
    axis.title.y = element_text(size = 16))


```

Now let's explore the spatial dimensions of this with more granularity by at this on an interactive map:

```{r}
#| eval: true
#| echo: true
#| output: true

cat_ex_map <- st_as_sf(ex_records, coords = c("Longitude", "Latitude"), crs= 4326)
cat_ex_map$cat_arrival <- factor(cat_ex_map$cat_arrival, levels = arrival_levels)

cat_ex_map %>%
  select(Class, Order, `Scientific name`, Extinction_extirpation, `Redlist category`,
         Prey_extirpation_year_earliest, Prey_extirpation_year_latest,
         Cat_arrival_year_earliest, Cat_arrival_year_latest, cat_arrival,
         Prey_quote) %>%
  mapview(zcol = "cat_arrival",
        col.regions = rev(plot_col),
        layer.name = "Extinction relative to arrival")

```

